---
title: "altair_scatterplot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{altair_scatterplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

When working in both R and Python, be sure to load reticulate, which helps with transferring between the two languages. You can also load r data sets to be use in Python.
```{r setup}
library(newprojdata303)
library(reticulate)
library(tidyverse)

```

By convention, always import pandas as pd. In python, "as" renames an object to something that is easier to type in order to save time for repeated use.

In the curly brackets at the top of the chunk, write "python" instead of R to indicate that this is Python code. Standard Python rules apply. 
```{python}
import pandas as pd
import altair as alt
```

```{python}
mpg = r.mpg

mpg
```

```{python}
alt.Chart(mpg).mark_point().encode(
    x='cty',
    y='hwy'
)
```


```{python}
alt.Chart(mpg).mark_point().encode(
    x='cty',
    y='hwy',
    color='manufacturer'
)
```

```{python}
alt.Chart(mpg).mark_point().encode(
    x='cty',
    y='hwy',
    color='manufacturer',
    tooltip=['manufacturer', 'model'] # show Name and Origin in a tooltip
).interactive()
```

```{python}
alt.Chart(mpg).mark_circle().encode(
    x='cty',
    y='hwy',
    color='manufacturer',
    tooltip=['manufacturer', 'model'] # show Name and Origin in a tooltip
).interactive()
```

