---
title: "df_apply"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{df_apply}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(newprojdata303)
```

## Description 

This function takes in a data frame, a function, and a filter. The function can 
be applied to a column of a data frame and return the column of the dataframe. 
The filter will return either a true or false.


## Use 

df_apply(.data, .fun, .filter, ...)

Our team used df_apply to round the lengths and widths of iris petals using 
the R preset iris data frame. (This would be useful when precision isn't needed
and the user wants to round to integers before analysis.) To use df_apply() to 
round numbers, simply use the premade function round() in the place of .fun, 
with is.numeric() as .filter and the data frame of your choice:

iris_function <- df_apply(iris, round, is.numeric)

The function can be used with any data set, mutate function, and filter 
function. The filter function will be applied to all columns, one by one, 
checking whether a condition is true. For instance, to change all text in a data 
frame to upper case, we can use the is.character() function as a filter, than 
apply the toupper() function to change columns of characters to upper case.

The df_apply() function returns an updated data frame.

## General requirements
Pass the df_apply function:

1. a data frame
2. a function which can be applied to a vector and change it
3. a function which can take a vector and check whether a condition is true for 
that vector (for instance, whether it contains a certain data type)


